3. Analyses Plan

### 2021-06-29 analysis/reports/00_reproduction.Rmd
Restructure the Project's Analysis

This document explores the data provided by @GACHTER2017 and finds that some of the variables are a little off. In addition, it fails to replicate the orginal article's figures. The following chapter will describe the available data. Subsequently, [data excerpts](#Tables) will demonstrate that the original _Gini coefficient_ varies within groups (in selected periods) and the _share of current endowment contributed_ is calculated using the last round's endowment. [Afterwards](#Downloads), the data is processsed and made available for downloads.The [Visualizations](#Visualizations) chapter will then present the original figures as well as figures reproduced using the original and re-calculated variables to demonstrate differences.

@@

GMTV <- read_dta(file="../data/gaechteretal/GMTV-data.dta") %>% data.table()

>>>>

This document explores the data provided by @GACHTER2017 and finds that some of the variables are a little off. In addition, it fails to replicate the orginal article's figures. The following chapter will describe the available data. Subsequently, [data excerpts](#Tables) will demonstrate that the original _Gini coefficient_ varies within groups (in selected periods) and the _share of current endowment contributed_ is calculated using the last round's endowment. [Afterwards](#Downloads), the data is processed and made available for downloads.The [Visualizations](#Visualizations) chapter will then present the original figures as well as figures reproduced using the original and re-calculated variables to demonstrate differences.

@@

GMTV <- read_dta(file="../../data/gaechteretal/GMTV-data.dta") %>% data.table()

### 2021-06-29 analysis/reports/rmd/01_dataGMTV.Rmd
Restructure the Project's Analysis

---
title: "1. Original Data"
description: |
  Reproducing Gächter et al.'s main findings.
author:
  - name: Hauke Roggenkamp 
    url: https://github.com/Howquez
    affiliation: CLICCS
    affiliation_url: https://www.cliccs.uni-hamburg.de/
date: "`r Sys.Date()`"
bibliography: biblio.bib
output:
  distill::distill_article:
    toc: true
    toc_depth: 2
    toc_float: false
    code_folding: true
---

@@

```{r readData}
DT <- read_dta(file="../data/gaechteretal/GMTV-data.dta") %>% data.table()
CT <- read_dta(file="../data/gaechteretal/GMTV-questionnaire-data.dta") %>% data.table()
```

@@

Within the appendix, one also learns that the NOPUNISH 10-period data were collected in sessions 5, 8 and 9. Subsetting the data correspondingly yields a ```data.table``` consisting of 23 observations and $4 \times 10 \times 23=$ `r NROW(noPunish10)` rows and 27 variables.

@@

- ```exp_num``` is a session identifier
- ```per``` denotes the period
- ```gr_id``` is a group identifier (carrying treatment information)
- ```subj_id``` is a subject identifier
- ```tokens``` reports a subject's endowment in a period
- ```other[1-3]``` report the other group members' endowments in a period
- ```gdp ``` equals the sum of endowments of a group in a period
- ```putin``` reports a subject's contribution in a period
- ```pu[1-3]``` report the other group members' contributions in a period
- ```sum``` equals the sum of contributions of a group in a period.
- ```gini``` reports a group's Gini coefficient in a period.
- ```mean``` reports the fraction of sum/lagged(gdp).

Additionally, I compute two new measures, ```gini2``` as well as ```share```. The new ```gini2``` differs from the original as it is constant within groups. ```share``` is equivalent to the original mean but relies on the current gdp instead of the past one as a denominator.

@@

# show table
noPunish10[order(gr_id, per),
           .(exp_num, gr_id, per, subj_id, # IDs
             tokens, other1, other2, other3, gdp, #endowments
             putin, pu1, pu2, pu3, sum, mean, share, # contributions
             gini, gini2, gini3
           )] %>% 
  head(n = 15) %>%
  kbl() %>% 
  # scroll_box(height = "200px") %>%
  kable_paper("hover", 
              full_width = TRUE,
              fixed_thead = TRUE)

@@

We see group 501 in the first two periods where subjects 511 to 514 start with an initial endowment of 20 which makes a GDP of 80. Subject 511 does not contribute (as represented in ```putin==0``` in the first row and ```pu1==0```in rows 2 to 4) while the others contribute 5, 15 and 15 in the first period, respectively. The ```sum``` of contributions therefore equals 35 period 1.

With a sum of contributions being 35 and a return of contributions of 1.5, each subject receives ```ceil(13.125)``` =14 points from the group project such that the next period's ```gdp``` equals 101.

The Gini coefficient (```gini```) is a little off as one can see in the second period: Even though this is a group level variable, the group has three different coefficient in the same period. With the same absolute amount of contributions as in the first period,the second period ```mean``` (based on the past ```gdp```) is larger than ```share``` in the second period, because the ```gdp``` grew.

@@

To resemble the oTree data, the original data has to be transformed a little. Therefore, some variables will be renamed, variables (such as ```stock```^[The income at the end of a period.] and ```gain```^[The tokens earned in a period.]) will be created and, most notably, the unit of observations will be groups such that subject-level variables will be dropped.

@@

# display data
GMTV %>%
  head(n = 15) %>%
  kbl() %>% 
  kable_paper("hover", 
              full_width = TRUE,
              fixed_thead = TRUE)

@@

```{r writeData}
save(GMTV, file = "../data/processed/rda/GMTV2017.rda")
write.csv(GMTV, file ="../data/processed/csv/GMTV2017.csv")

@@

```{r downloadData}
GMTV %>%
      download_this(
        output_name = "GMTV2017noPunish10",
        output_extension = ".csv", # CSV output
        button_label = "Download csv",
        button_type = "default"
        )
```

@@

# Tables

```{r regressions}
# reg gdp punish if longgame==0  & treat<3 & per==10, cluster(gr_id)
lm(gdp ~ punish, data = DT[longgame==0 & treat < 3 & per==10])
```

@@

Even though the figures above were created using the original data, they do not look the same as the original figures. This becomes obvious if one looks at the figures that visualize the contributions. The arising question therefore is whether I made a mistake in processing and interpreting the data. After all I thought to have found wrong Gini coefficients and struggle to interpret the ```mean``` of the original data.

@@

# save data
fileName <- "GMTV2017"
save(GMTVFirstRound, 
     file = paste0("../data/processed/rda/", fileName, "_R1", ".rda"))
write.csv(GMTVFirstRound, 
          file = paste0("../data/processed/csv/", fileName, "_R1", ".csv"))
```

`r NROW(GMTVFirstRound)` participants yield `r NROW(GMTVFirstRound)` independent observations that are stored in a data table called ```GMTVFirstRound```. These observations are displayed below and can, once more, be downloaded by a click on the button. As before, the data is also saved in ```../data/processed/```.

```{r displayFirstRound}
# display data
GMTVFirstRound %>%
  head(n = 15) %>%
  kbl() %>% 
  # scroll_box(height = "200px") %>%
  kable_paper("hover", 
              full_width = TRUE,
              fixed_thead = TRUE)

# create download button
GMTVFirstRound %>%
      download_this(
        output_name = "GMTV2017_R1",
        output_extension = ".csv", # CSV output
        button_label = "Download csv",
        button_type = "default"
        )
```

@@

# write data
fileName <- "GMTV2017"
save(GMTVCovariates, 
     file = paste0("../data/processed/rda/", fileName, "_COVS", ".rda"))
write.csv(GMTVCovariates, 
          file = paste0("../data/processed/csv/", fileName, "_COVS", ".csv"))

>>>>

---
title: "1. Original Data"
description: |
  Reproducing Gächter et al.'s main findings.
author:
  - name: Hauke Roggenkamp 
    url: https://github.com/Howquez
    affiliation: CLICCS
    affiliation_url: https://www.cliccs.uni-hamburg.de/
date: "`r Sys.Date()`"
bibliography: ../biblio.bib
output:
  distill::distill_article:
    toc: true
    toc_depth: 2
    toc_float: false
    code_folding: true
---

@@

```{r readData}
DT <- read_dta(file="../../../data/gaechteretal/GMTV-data.dta") %>% data.table()
CT <- read_dta(file="../../../data/gaechteretal/GMTV-questionnaire-data.dta") %>% data.table()
```

@@

Within the appendix, one also learns that the NOPUNISH 10-period data were collected in sessions 5, 8 and 9. Subsetting the data correspondingly yields a `data.table` consisting of 23 observations and $4 \times 10 \times 23=$ `r NROW(noPunish10)` rows and 27 variables.

@@

-   `exp_num` is a session identifier
-   `per` denotes the period
-   `gr_id` is a group identifier (carrying treatment information)
-   `subj_id` is a subject identifier
-   `tokens` reports a subject's endowment in a period
-   `other[1-3]` report the other group members' endowments in a period
-   `gdp` equals the sum of endowments of a group in a period
-   `putin` reports a subject's contribution in a period
-   `pu[1-3]` report the other group members' contributions in a period
-   `sum` equals the sum of contributions of a group in a period.
-   `gini` reports a group's Gini coefficient in a period.
-   `mean` reports the fraction of sum/lagged(gdp).

Additionally, I compute two new measures, `gini2` as well as `share`. The new `gini2` differs from the original as it is constant within groups. `share` is equivalent to the original mean but relies on the current gdp instead of the past one as a denominator.

@@

if(knitr::is_html_output()){
  # show table
  noPunish10[order(gr_id, per),
             .(exp_num, gr_id, per, subj_id, # IDs
               tokens, other1, other2, other3, gdp, #endowments
               putin, pu1, pu2, pu3, sum, mean, share, # contributions
               gini, gini2, gini3
             )] %>% 
    head(n = 15) %>%
    kbl() %>% 
    # scroll_box(height = "200px") %>%
    kable_paper("hover", 
                full_width = TRUE,
                fixed_thead = TRUE)
  }

@@

We see group 501 in the first two periods where subjects 511 to 514 start with an initial endowment of 20 which makes a GDP of 80. Subject 511 does not contribute (as represented in `putin==0` in the first row and `pu1==0`in rows 2 to 4) while the others contribute 5, 15 and 15 in the first period, respectively. The `sum` of contributions therefore equals 35 period 1.

With a sum of contributions being 35 and a return of contributions of 1.5, each subject receives `ceil(13.125)` =14 points from the group project such that the next period's `gdp` equals 101.

The Gini coefficient (`gini`) is a little off as one can see in the second period: Even though this is a group level variable, the group has three different coefficient in the same period. With the same absolute amount of contributions as in the first period,the second period `mean` (based on the past `gdp`) is larger than `share` in the second period, because the `gdp` grew.

@@

To resemble the oTree data, the original data has to be transformed a little. Therefore, some variables will be renamed, variables (such as `stock`^[The income at the end of a period.] and `gain`^[The tokens earned in a period.]) will be created and, most notably, the unit of observations will be groups such that subject-level variables will be dropped.

@@

if(knitr::is_html_output()){
  # display data
  GMTV %>%
    head(n = 15) %>%
    kbl() %>% 
    kable_paper("hover", 
                full_width = TRUE,
                fixed_thead = TRUE)
  }

@@

```{r writeData}
save(GMTV, file = "../../../data/processed/rda/GMTV2017.rda")
write.csv(GMTV, file ="../../../data/processed/csv/GMTV2017.csv")

@@

```{r downloadData}
if(knitr::is_html_output()){
  GMTV %>%
        download_this(
          output_name = "GMTV2017noPunish10",
          output_extension = ".csv", # CSV output
          button_label = "Download csv",
          button_type = "default"
        )
  }
```

@@

```{r regressions, eval = FALSE}
# reg gdp punish if longgame==0  & treat<3 & per==10, cluster(gr_id)
lm(gdp ~ punish, data = DT[longgame==0 & treat < 3 & per==10])
```

@@

Even though the figures above were created using the original data, they do not look the same as the original figures. This becomes obvious if one looks at the figures that visualize the contributions.

@@

# save data
fileName <- "GMTV2017"
save(GMTVFirstRound, 
     file = paste0("../../../data/processed/rda/", fileName, "_R1", ".rda"))
write.csv(GMTVFirstRound, 
          file = paste0("../../../data/processed/csv/", fileName, "_R1", ".csv"))
```

`r NROW(GMTVFirstRound)` participants yield `r NROW(GMTVFirstRound)` independent observations that are stored in a data table called `GMTVFirstRound`. These observations are displayed below and can, once more, be downloaded by a click on the button (HTML-only feature). As before, the data is also saved in `../../../data/processed/`.

```{r displayFirstRound}

if(knitr::is_html_output()){
  # display data
  GMTVFirstRound %>%
    head(n = 15) %>%
    kbl() %>% 
    # scroll_box(height = "200px") %>%
    kable_paper("hover", 
                full_width = TRUE,
                fixed_thead = TRUE)

  # create download button
  GMTVFirstRound %>%
        download_this(
          output_name = "GMTV2017_R1",
          output_extension = ".csv", # CSV output
          button_label = "Download csv",
          button_type = "default"
        )
  }
```

@@

# write data
fileName <- "GMTV2017"
save(GMTVCovariates, 
     file = paste0("../../../data/processed/rda/", fileName, "_COVS", ".rda"))
write.csv(GMTVCovariates, 
          file = paste0("../../../data/processed/csv/", fileName, "_COVS", ".csv"))

### 2021-06-29 analysis/reports/rmd/02_dataProcessing.Rmd
Restructure the Project's Analysis

---
title: "2. Simulated Data"
description: |
  Making the raw oTree output tangible.
author:
  - name: Hauke Roggenkamp 
    url: https://github.com/Howquez
    affiliation: CLICCS
    affiliation_url: https://www.cliccs.uni-hamburg.de/
date: "`r Sys.Date()`"
bibliography: biblio.bib
output:
  distill::distill_article:
    toc: true
    toc_depth: 2
    toc_float: false
    code_folding: true
---

@@

# Simulations

```{r readData}
# read simulated data
DT <- read.csv("../data/simulation/all_apps_wide.csv", stringsAsFactors = FALSE) %>%
  data.table()
```

I simulated `r NROW(DT)` participants. To do so, I wrote oTree bots that follow rather simple rules and act mostly random. As a consequence, this data yields no information apart from the structure of the data itself.

Eventually, this report can also be used to process real data. This only requires us to read a different data set.^[It may also require us to merge multiple data sets -- that depends on whether the lab delivers one or several data sets and remains to be discussed.]

@@

# save data
fileName <- "replication2021"
save(replicationFirstRound, file = paste0("../data/processed/rda/", fileName, "_R1", ".rda"))
write.csv(replicationFirstRound, file = paste0("../data/processed/csv/", fileName, "_R1", ".csv"))

# display data
replicationFirstRound %>%
  head(n = 12) %>%
  kbl() %>% 
  # scroll_box(height = "200px") %>%
  kable_paper("hover", 
              full_width = TRUE,
              fixed_thead = TRUE)

# create download button
replicationFirstRound %>%
      download_this(
        output_name = glue("{Sys.Date()}_Simulation_firstRound"),
        output_extension = ".csv", # CSV output
        button_label = "Download csv",
        button_type = "default"
        )
```

The data is saved in two formats (```csv``` and ```rda```) in ```../data/processed/```. If you have not downloaded the repository and access the html file, you can also obtain a csv-file by clicking on the button above.

@@

# display data
subset %>%
  head(n = 12) %>%
  kbl() %>% 
  scroll_box(width = "100%") %>%
  kable_paper("hover", 
              full_width = TRUE,
              fixed_thead = TRUE)
```

@@

The result contains ```NROW(DT)/4``` observations in 10 rounds and is saved as a ```csv``` and as a ```rda``` file in ```../data/processed/```.

```{r saveData}
save(replication, file = paste0("../data/processed/rda/", fileName, ".rda"))
write.csv(replication, file = paste0("../data/processed/csv/", fileName, ".csv"))
```

@@

```{r displayData}
replication %>%
  head(n = 12) %>%
  kbl() %>% 
  # scroll_box(height = "200px") %>%
  kable_paper("hover", 
              full_width = TRUE,
              fixed_thead = TRUE)
```

```{r downloadData}
replication %>%
      download_this(
        output_name = glue("{Sys.Date()}_Simulation"),
        output_extension = ".csv", # CSV output
        button_label = "Download csv",
        button_type = "default"
        )
```

@@

# write data
fileName <- "replication2021"
save(replicationCovariates, 
     file = paste0("../data/processed/rda/", fileName, "_COVS", ".rda"))
write.csv(replicationCovariates, 
          file = paste0("../data/processed/csv/", fileName, "_COVS", ".csv"))

>>>>

---
title: "2. Simulated Data"
description: |
  Making the raw oTree output tangible.
author:
  - name: Hauke Roggenkamp 
    url: https://github.com/Howquez
    affiliation: CLICCS
    affiliation_url: https://www.cliccs.uni-hamburg.de/
date: "`r Sys.Date()`"
bibliography: ../biblio.bib
output:
  distill::distill_article:
    toc: true
    toc_depth: 2
    toc_float: false
    code_folding: true
---

@@

# Replication Data

```{r readData}
# read replication data
DT <- read.csv(paste("../..", recentCSV, sep = "/"), 
               stringsAsFactors = FALSE) %>%
  data.table()
```

@@

# save data
fileName <- "replication2021"
save(replicationFirstRound, file = paste0("../../../data/processed/rda/", fileName, "_R1", ".rda"))
write.csv(replicationFirstRound, file = paste0("../../../data/processed/csv/", fileName, "_R1", ".csv"))

if(knitr::is_html_output()){
  # display data
  replicationFirstRound %>%
    head(n = 12) %>%
    kbl() %>% 
    # scroll_box(height = "200px") %>%
    kable_paper("hover", 
                full_width = TRUE,
                fixed_thead = TRUE)

  # create download button
  replicationFirstRound %>%
        download_this(
          output_name = glue("{Sys.Date()}_Simulation_firstRound"),
          output_extension = ".csv", # CSV output
          button_label = "Download csv",
          button_type = "default"
        )
  }
```

The data is saved in two formats (```csv``` and ```rda```) in ```../../../data/processed/```. If you have not downloaded the repository and access the html file, you can also obtain a csv-file by clicking on the button above (HTML-only feature).

@@

if(knitr::is_html_output()){
  # display data
  subset %>%
    head(n = 12) %>%
    kbl() %>% 
    scroll_box(width = "100%") %>%
    kable_paper("hover", 
                full_width = TRUE,
                fixed_thead = TRUE)
  }
```

@@

The result contains ```NROW(DT)/4``` observations in 10 rounds and is saved as a ```csv``` and as a ```rda``` file in ```../../../data/processed/```.

```{r saveData}
save(replication, file = paste0("../../../data/processed/rda/", fileName, ".rda"))
write.csv(replication, file = paste0("../../../data/processed/csv/", fileName, ".csv"))
```

@@

```{r displayData}

if(knitr::is_html_output()){
  replication %>%
    head(n = 12) %>%
    kbl() %>% 
    # scroll_box(height = "200px") %>%
    kable_paper("hover", 
                full_width = TRUE,
                fixed_thead = TRUE)
  }
```

```{r downloadData}
if(knitr::is_html_output()){
  replication %>%
        download_this(
          output_name = glue("{Sys.Date()}_Simulation"),
          output_extension = ".csv", # CSV output
          button_label = "Download csv",
          button_type = "default"
        )
  }
```

@@

# write data
fileName <- "replication2021"
save(replicationCovariates, 
     file = paste0("../../../data/processed/rda/", fileName, "_COVS", ".rda"))
write.csv(replicationCovariates, 
          file = paste0("../../../data/processed/csv/", fileName, "_COVS", ".csv"))

### 2021-06-29 analysis/reports/rmd/03_dataAnalyses.Rmd
Restructure the Project's Analysis

---
title: "3. Analyses Plan"
description: |
  Analyze and compare data
author:
  - name: Hauke Roggenkamp 
    url: https://github.com/Howquez
    affiliation: CLICCS
    affiliation_url: https://www.cliccs.uni-hamburg.de/
date: "`r Sys.Date()`"
bibliography: biblio.bib
output:
  distill::distill_article:
    toc: true
    toc_depth: 2
    toc_float: false
    code_folding: true
---

@@

```{r readData}
base::load(file="../data/processed/rda/GMTV2017.rda")              # read GMTV
base::load(file="../data/processed/rda/GMTV2017_R1.rda")           # read GMTVFirstRound
base::load(file="../data/processed/rda/GMTV2017_COVS.rda")         # read GMTVCovariates
base::load(file="../data/processed/rda/replication2021.rda")       # read replication
base::load(file="../data/processed/rda/replication2021_R1.rda")    # read replicationFirstRound
base::load(file="../data/processed/rda/replication2021_COVS.rda")  # read replicationCovariates
```

@@

rs1 <- wilcox.test(replication, GMTV)$p.value %>% round(digits = 4) %>% formatC(format = "f", digits = 4)

@@

rs2 <- wilcox.test(replication, GMTV)$p.value %>% round(digits = 4) %>% formatC(format = "f", digits = 4)

>>>>

---
title: "3. Analyses Plan"
description: |
  Analyze and compare data
author:
  - name: Hauke Roggenkamp 
    url: https://github.com/Howquez
    affiliation: CLICCS
    affiliation_url: https://www.cliccs.uni-hamburg.de/
date: "`r Sys.Date()`"
bibliography: ../biblio.bib
output:
  distill::distill_article:
    toc: true
    toc_depth: 2
    toc_float: false
    code_folding: true
---

@@

```{r readData}
base::load(file="../../../data/processed/rda/GMTV2017.rda")              # read GMTV
base::load(file="../../../data/processed/rda/GMTV2017_R1.rda")           # read GMTVFirstRound
base::load(file="../../../data/processed/rda/GMTV2017_COVS.rda")         # read GMTVCovariates
base::load(file="../../../data/processed/rda/replication2021.rda")       # read replication
base::load(file="../../../data/processed/rda/replication2021_R1.rda")    # read replicationFirstRound
base::load(file="../../../data/processed/rda/replication2021_COVS.rda")  # read replicationCovariates
```

@@

rs1 <- wilcox.test(replication, 
                   GMTV,
                   exact = FALSE)$p.value %>% 
  round(digits = 4) %>% 
  formatC(format = "f", 
          digits = 4)

@@

rs2 <- wilcox.test(replication, 
                   GMTV,
                   exact = FALSE)$p.value %>% 
  round(digits = 4) %>% 
  formatC(format = "f", 
          digits = 4)
